Pre-/Post-Trikafta in CFRD

### 2024-07-17 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.qmd
updates for Colleen Wood finished

---
title: "Pre-/Post-Trikafta in CFRD"
author: "Tim Vigers"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(gtsummary)
library(arsenal)
library(rspiro)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r}
load("./Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
# Baseline HbA1c, BMI group pre- and post-modulator
df <- df %>%
  group_by(`CFF ID`) %>%
  mutate(
    `Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"],
    `Baseline BMI` = `BMI Percentile`[`Pre or Post` == "Pre"],
    `Post BMI` = `BMI Percentile`[`Pre or Post` == "Post"]
  ) %>%
  ungroup()
df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf), right = F)
df$`Baseline BMI` <- cut(df$`Baseline BMI`, c(-Inf, 85, Inf), right = F)
df$`Post BMI` <- cut(df$`Post BMI`, c(-Inf, 85, Inf), right = F)
# Pre-modulator BMI percentile
df <- df %>%
  group_by(`CFF ID`) %>%
  mutate(`Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"]) %>%
  ungroup()
df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf), right = F)
# Race neutral equations
df$ppFEV1gl <- pctpred_GLIgl(
  age = df$Age, height = df$Height / 100, gender = df$Sex, FEV1 = df$FEV1
)
df$ppFVCgl <- pctpred_GLIgl(
  age = df$Age, height = df$Height / 100, gender = df$Sex, FVC = df$FEV1
)
# Columns formats
df$`Num. A1c Values` <- as.numeric(df$`Num. A1c Values`)
```

@@

```{r}
df %>%
  filter(`Pre or Post` == "Pre") %>%
  select(
    Age, `BMI Percentile`, Weight, Sex, ppFEV1, ppFEV1gl, ppFVC,
    ppFVCgl, `G-tube in past 12 months`, `Pancreatic Status`, Site,
    Genotype, `Age Group`, Race
  ) %>%
  tbl_summary(
    statistic = list(c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ "{mean} ± {sd}"),
    missing_text = "Missing"
  )
```

@@

```{r}
cgm <- df[!is.na(df$average_sensor), ]
cgm %>%
  filter(`Pre or Post` == "Pre") %>%
  select(
    Age, `BMI Percentile`, Weight, Sex, ppFEV1, ppFEV1gl, ppFVC,
    ppFVCgl, `G-tube in past 12 months`, `Pancreatic Status`, Site,
    Genotype, `Age Group`, Race
  ) %>%
  tbl_summary(
    type = list(
      c(
        "Age", "BMI Percentile", "Weight", "ppFEV1", "ppFEV1gl", "ppFVC",
        "ppFVCgl"
      ) ~ "continuous2"
    ),
    statistic = list(all_continuous2() ~
      c("{median} [{p25}, {p75}]", "{min}, {max}")),
    digits = list(everything() ~ 2)
  )
```

>>>>

---
title: "Pre-/Post-Trikafta in CFRD"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /home/timvigers/Documents/Miscellaneous/zotero.bib
csl: /home/timvigers/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(gtsummary)
library(arsenal)
library(rspiro)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Colleen Wood"
)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| include: false
# # Import CHCO data
# chco_bmi_pft <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/chco_bmi_pft.csv",
#   na.strings = "", check.names = F
# )
# chco_bmi_pft$Date <- mdy(chco_bmi_pft$Date)
# chco_cgm <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/chco_cgm.csv",
#   na.strings = "", check.names = F
# )
# chco_cgm$Date <- date(parse_date_time(chco_cgm$date_cgm_placement,
#   orders = c("mdyHM", "mdy")
# ))
# chco_cgm$`CFF ID` <- as.numeric(sub("_.*", "", chco_cgm$subject_id))
# chco_cgm$Visit <- sub(".*_", "", chco_cgm$subject_id)
# chco_glycemic <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/chco_glycemic.csv",
#   na.strings = "", check.names = F
# )
# chco_glycemic$Date <- mdy(chco_glycemic$Date)
# chco_registry <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/chco_registry.csv",
#   na.strings = "", check.names = F
# )
# chco_registry$Start <- NULL
# # Import Montana
# montana_bmi_pft <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/montana_bmi_pft.csv",
#   na.strings = "", check.names = F
# )
# montana_bmi_pft$Date <- mdy(montana_bmi_pft$Date)
# montana_cgm <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/montana_cgm.csv",
#   na.strings = "", check.names = F
# )
# montana_cgm$Date <- date(parse_date_time(montana_cgm$date_cgm_placement,
#   orders = c("mdyHM", "mdy")
# ))
# montana_cgm$`CFF ID` <- as.numeric(sub("_.*", "", montana_cgm$subject_id))
# montana_cgm$Visit <- sub(".*_", "", montana_cgm$subject_id)
# montana_cgm$Date <- mdy(montana_cgm$Date)
# montana_demo <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/montana_demo.csv",
#   na.strings = "", check.names = F
# )
# montana_genotypes <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/montana_genotypes.csv",
#   na.strings = "", check.names = F
# )
# montana_gtubes <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/montana_gtubes.csv",
#   na.strings = "", check.names = F
# )
# montana_gtubes$Date <- NULL
# montana_pancreatic <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/montana_pancreatic.csv",
#   na.strings = "", check.names = F
# )
# montana_pancreatic$Date <- NULL
# montana_pancreatic$`Pancreatic Status` <- "Insufficient"
# montana_glycemic_registry <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/montana_glycemic_registry.csv",
#   na.strings = "", check.names = F
# )
# montana_glycemic_registry$Date <- mdy(montana_glycemic_registry$Date)
# # Import UW
# uw_registry <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/uw_registry.csv",
#   na.strings = "", check.names = F
# )
# uw_registry <- uw_registry %>% rename(Date = Start)
# uw_registry$Date <- mdy(uw_registry$Date)
# uw_bmi <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/uw_bmi.csv",
#   na.strings = "", check.names = F
# )
# uw_bmi$Date <- mdy(uw_bmi$Date)
# uw_pft <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/uw_pft.csv",
#   na.strings = "", check.names = F
# )
# uw_pft$Date <- mdy(uw_pft$Date)
# uw_cgm <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/uw_cgm.csv",
#   na.strings = "", check.names = F
# )
# uw_cgm$`CFF ID` <- as.numeric(sub("_.*", "", uw_cgm$subject_id))
# uw_cgm$Visit <- sub(".*_", "", uw_cgm$subject_id)
# uw_demo <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/uw_demo.csv",
#   na.strings = "", check.names = F
# )
# uw_genotypes <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/uw_registry.csv",
#   na.strings = "", check.names = F
# )
# uw_glycemic <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/uw_glycemic.csv",
#   na.strings = "", check.names = F
# )
# uw_glycemic$Date <- mdy(uw_glycemic$Date)
# uw_gtubes <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/uw_gtubes.csv",
#   na.strings = "", check.names = F
# )
# uw_gtubes$`Date of Encounter` <- NULL
# uw_pancreatic <- read.csv(
#   "./Prepost Trikafta/Data_Cleaned/uw_pancreatic.csv",
#   na.strings = "", check.names = F
# )
# uw_pancreatic$Date <- NULL
# # Format and merge
# # CHCO
# chco_long <- full_join(chco_glycemic, chco_bmi_pft)
# chco_long <- full_join(chco_registry, chco_long)
# chco_long <- full_join(chco_long, chco_cgm)
# chco_long$site <- "CHCO"
# chco_long$`Hispanic/Latinx`[chco_long$`Hispanic/Latinx`] <- "Yes"
# chco_long$`Hispanic/Latinx`[chco_long$`Hispanic/Latinx` == "FALSE"] <- "No"
# # Montana
# montana_registry <- full_join(montana_demo, montana_genotypes)
# montana_registry <- full_join(montana_registry, montana_pancreatic)
# montana_registry <- full_join(montana_registry, montana_gtubes)
# montana_long <- full_join(montana_glycemic_registry, montana_bmi_pft)
# montana_long <- full_join(montana_registry, montana_long)
# montana_long <- full_join(montana_long, montana_cgm)
# montana_long <- montana_long %>% rename(HbA1c_original = HbA1c)
# montana_long <- montana_long %>% mutate(
#   hba1c_spc = str_detect(HbA1c_original, " "),
#   HbA1c = as.numeric(ifelse(hba1c_spc, sub(" .*", "", montana_long$HbA1c_original), HbA1c_original))
# )
# montana_long <- montana_long %>% select(-c(hba1c_spc, HbA1c_original))
# montana_long$site <- "Montana"
# montana_long <- montana_long %>%
#   rename(`Hispanic/Latinx` = `Is Patient of Hispanic Origin`)
# # UW
# uw_bmi_pft <- full_join(uw_bmi, uw_pft)
# uw_registry <- full_join(uw_registry, uw_demo)
# uw_registry <- full_join(uw_registry, uw_gtubes)
# uw_long <- full_join(uw_glycemic, uw_bmi_pft)
# uw_long <- full_join(uw_registry, uw_long)
# uw_long <- full_join(uw_long, uw_cgm)
# uw_long$site <- "UW"
# uw_long <- uw_long %>%
#   rename(`Hispanic/Latinx` = `Is Patient of Hispanic Origin`)
# # Put everything together
# analysis_data <- full_join(chco_long, montana_long)
# analysis_data <- full_join(analysis_data, uw_long)
# # Race
# analysis_data$Race <- factor(analysis_data$Race,
#   levels = c(
#     "American Indian or Alaska Native", "Black", "Black or African American",
#     "Caucasian", "Other", "Some other race", "Two or more races", "Unknown",
#     "White"
#   ),
#   labels = c(
#     "Other", "Black", "Black", "White", "Other", "Other", "Other",
#     "Unknown", "White"
#   )
# )
# df <- analysis_data
load("/home/timvigers/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
# Ethnicity
chco_demo <- read.csv("/home/timvigers/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Raw/Data From CHCO/Registry/Copy of Earliest Trikafta Dates in Registry_updated 2.2.21 cc modified_cfrdyn.csv")
chco_demo$Hispanic <- NA
chco_demo$Hispanic[chco_demo$Hispanic.Latinx] <- "Yes"
chco_demo$Hispanic[!chco_demo$Hispanic.Latinx] <- "No"
chco_demo <- chco_demo[, c("CFF.ID", "Hispanic")]

montana_demo <- read.csv("/home/timvigers/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/montana_demo.csv")
montana_demo$Hispanic <- montana_demo$Is.Patient.of.Hispanic.Origin
montana_demo <- montana_demo[, c("CFF.ID", "Hispanic")]

uw_demo <- read.csv("/home/timvigers/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/uw_demo.csv")
uw_demo$Hispanic <- uw_demo$Is.Patient.of.Hispanic.Origin
uw_demo <- uw_demo[, c("CFF.ID", "Hispanic")]

eth <- do.call(rbind, list(chco_demo, montana_demo, uw_demo))
colnames(eth) <- c("CFF ID", "Hispanic")

df <- left_join(df, eth)
```

```{r}
# Baseline HbA1c, BMI group pre- and post-modulator
df <- df %>%
  group_by(`CFF ID`) %>%
  mutate(
    `Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"],
    `Baseline BMI` = `BMI Percentile`[`Pre or Post` == "Pre"],
    `Post BMI` = `BMI Percentile`[`Pre or Post` == "Post"]
  ) %>%
  ungroup()
df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf),
  right = F
)
df$`Baseline BMI` <- cut(df$`Baseline BMI`, c(-Inf, 85, Inf), right = F)
df$`Post BMI` <- cut(df$`Post BMI`, c(-Inf, 85, Inf), right = F)
# Pre-modulator BMI percentile
df <- df %>%
  group_by(`CFF ID`) %>%
  mutate(`Baseline HbA1c` = HbA1c[`Pre or Post` == "Pre"]) %>%
  ungroup()
df$`Baseline HbA1c` <- cut(df$`Baseline HbA1c`, c(-Inf, 6.5, Inf), right = F)
# Race neutral equations
df$ppFEV1gl <- pctpred_GLIgl(
  age = df$Age, height = df$Height / 100, gender = df$Sex, FEV1 = df$FEV1
)
df$ppFVCgl <- pctpred_GLIgl(
  age = df$Age, height = df$Height / 100, gender = df$Sex, FVC = df$FEV1
)
# Columns formats
df$`Num. A1c Values` <- as.numeric(df$`Num. A1c Values`)
```

@@

```{r}
df %>%
  filter(`Pre or Post` == "Pre") %>%
  select(
    Age, `BMI Percentile`, Weight, Sex, ppFEV1, ppFEV1gl, ppFVC,
    ppFVCgl, `G-tube in past 12 months`, `Pancreatic Status`, Site,
    Genotype, `Age Group`, Race,Hispanic
  ) %>%
  tbl_summary(
    statistic = list(c("ppFEV1", "ppFEV1gl", "ppFVC", "ppFVCgl") ~ 
                       "{mean} ± {sd}"),
    missing_text = "Missing"
  )
```

@@

```{r}
cgm <- df[!is.na(df$mean_sensor), ]
cgm %>%
  filter(`Pre or Post` == "Pre") %>%
  select(
    Age, `BMI Percentile`, Weight, Sex, ppFEV1, ppFEV1gl, ppFVC,
    ppFVCgl, `G-tube in past 12 months`, `Pancreatic Status`, Site,
    Genotype, `Age Group`, Race
  ) %>%
  tbl_summary(
    type = list(
      c(
        "Age", "BMI Percentile", "Weight", "ppFEV1", "ppFEV1gl", "ppFVC",
        "ppFVCgl"
      ) ~ "continuous2"
    ),
    statistic = list(all_continuous2() ~
      c("{median} [{p25}, {p75}]", "{min}, {max}")),
    digits = list(everything() ~ 2)
  )
```

### 2024-07-17 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_merge.Rmd
updates for Colleen Wood finished

NOTE: I HAVE THE SHARED DRIVE AS S:, to change if mapped to another letter
```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(lubridate)
library(stringr)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/Documents/Work/Vigers/CF/Colleen Wood"
}
knitr::opts_knit$set(root.dir = home_dir)
```

@@

# bmi, pulm function 577 obs
# 8 vars MRN - ID, date for visits; start - ? start of trt?, fev, bmi, height, weight, fec 
chco_bmi_pft = read_csv("./Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv")
chco_bmi_pft$Date = mdy(chco_bmi_pft$Date) 
# continuous glucose monitoring, 23 obs
# 105 vars subjectid - number_pre/post for visit and a date; numbers dont seem to correspond to mrn in bmi pft
chco_cgm = read_csv("./Prepost Triakfta/Data_Cleaned/chco_cgm.csv")
chco_cgm$Date = date(parse_date_time(chco_cgm$date_cgm_placement,
                                                orders = c("mdyHM","mdy")))
# chco glycemic, 261 obs
# 7 var MRN and CFF id (=subjectid?), date for visits; start... etc
chco_glycemic = read_csv("./Prepost Triakfta/Data_Cleaned/chco_glycemic.csv")
chco_glycemic$Date = mdy(chco_glycemic$Date)
# chco registry 108 obs
# CFF ID and MRN, start date
chco_registry = read_csv("./Prepost Triakfta/Data_Cleaned/chco_registry.csv")
###################################

@@

# NEED TO RENAME VARIABLES TO MATCH CHCO#
# bmi, pulm function 975 obs
montana_bmi_pft <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_bmi_pft.csv")

# cgm 16 obs
montana_cgm <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_cgm.csv")
montana_cgm$Date = date(parse_date_time(montana_cgm$date_cgm_placement,
                                                orders = c("mdyHM","mdy")))

# registry: we need to combine demo (126), genotypes (126), pancreatic (241), gtubes (155)
montana_demo <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_demo.csv")
montana_genotypes <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_genotypes.csv")
montana_gtubes <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_gtubes.csv") # what are the dates here
montana_pancreatic <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_pancreatic.csv") # what is this? just dates and cffid
# ask tim:what do we need to do with gtubes and pancreatic dates? dont look like chco data

# glycemic 279
montana_glycemic_registry <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv")
montana_glycemic_registry$Date = mdy(montana_glycemic_registry$Date)
# looks identical
###################################

###### IMPORT UW DATA #############
uw_registry <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_registry.csv") #cffid and start
uw_bmi <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_bmi.csv") # date, cffid, height, bmi (height bmi need rename)
uw_pft <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_pft.csv") # datee, cffid, fev, fvc need renaming
uw_cgm <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_cgm.csv") # same code as montana
uw_demo <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_demo.csv") 
uw_genotypes <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_registry.csv") # no genotype data here?
uw_glycemic <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_glycemic.csv") # looks good, same as chco
uw_gtubes <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_gtubes.csv") # rename date of encounter to date
uw_pancreatic <- read_csv("./Prepost Triakfta/Data_Cleaned/uw_pancreatic.csv") # same as montana, no pancreatic data though

@@

###### IMPORT UM DATA ########
#um_all <- read_csv("./Prepost Triakfta/Data_Cleaned/um_all.csv")
###################################

@@

# keep original data, add in a new one to match CHCO
montana_demo = montana_demo %>% rename(race_original = Race)
montana_demo = montana_demo %>% mutate(Race = case_when(race_original == "White" ~ "Caucasian",
                                                        race_original == "Black or African American" ~ "Black",
                                                        race_original == NA ~ "Unknown",
                                                        TRUE ~ "Other"))
# rename the bmi_pft to chco names
montana_bmi_pft = montana_bmi_pft %>% rename(FVC = `Measure of FVC`,
                                             FEV1 = `Measure of FEV1`,
                                             Height = `Patients height`,
                                             Weight = `Weight of Patient`)
montana_bmi_pft$Date = mdy(montana_bmi_pft$Date)

@@

# keep original data, add in a new one to match CHCO
uw_demo = uw_demo %>% rename(race_original = Race)
uw_demo = uw_demo %>% mutate(Race = case_when(race_original == "White" ~ "Caucasian",
                                                        race_original == "Black or African American" ~ "Black",
                                                        race_original == NA ~ "Unknown",
                                                        TRUE ~ "Other"))
# make bmi pft
uw_bmi_pft = full_join(uw_bmi, uw_pft)
# rename the bmi_pft to chco names
uw_bmi_pft = uw_bmi_pft %>% rename(FVC = `Measure of FVC`,
                                   FEV1 = `Measure of FEV1`,
                                   Height = `Patients height`,
                                   BMI = `BMI Value`)

@@

# export csv analysis_data
write.csv(analysis_data, file = "./Prepost Triakfta/Data_Cleaned/analysis_data.csv")
```

>>>>

NOTE: I HAVE THE SHARED DRIVE AS S:, to change if mapped to another letter
```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(lubridate)
library(stringr)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Colleen Wood"
)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

# bmi, pulm function 577 obs
# 8 vars MRN - ID, date for visits; start - ? start of trt?, fev, bmi, height, weight, fec 
chco_bmi_pft = read_csv("./Prepost Trikafta/Data_Cleaned/chco_bmi_pft.csv")
chco_bmi_pft$Date = mdy(chco_bmi_pft$Date) 
# continuous glucose monitoring, 23 obs
# 105 vars subjectid - number_pre/post for visit and a date; numbers dont seem to correspond to mrn in bmi pft
chco_cgm = read_csv("./Prepost Trikafta/Data_Cleaned/chco_cgm.csv")
chco_cgm$Date = date(parse_date_time(chco_cgm$date_cgm_placement,
                                                orders = c("mdyHM","mdy")))
# chco glycemic, 261 obs
# 7 var MRN and CFF id (=subjectid?), date for visits; start... etc
chco_glycemic = read_csv("./Prepost Trikafta/Data_Cleaned/chco_glycemic.csv")
chco_glycemic$Date = mdy(chco_glycemic$Date)
# chco registry 108 obs
# CFF ID and MRN, start date
chco_registry = read_csv("./Prepost Trikafta/Data_Cleaned/chco_registry.csv")
###################################

@@

# NEED TO RENAME VARIABLES TO MATCH CHCO#
# bmi, pulm function 975 obs
montana_bmi_pft <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_bmi_pft.csv")

# cgm 16 obs
montana_cgm <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_cgm.csv")
montana_cgm$Date = date(parse_date_time(montana_cgm$date_cgm_placement,
                                                orders = c("mdyHM","mdy")))

# registry: we need to combine demo (126), genotypes (126), pancreatic (241), gtubes (155)
montana_demo <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_demo.csv")
montana_genotypes <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_genotypes.csv")
montana_gtubes <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_gtubes.csv") # what are the dates here
montana_pancreatic <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_pancreatic.csv") # what is this? just dates and cffid
# ask tim:what do we need to do with gtubes and pancreatic dates? dont look like chco data

# glycemic 279
montana_glycemic_registry <- read_csv("./Prepost Trikafta/Data_Cleaned/montana_glycemic_registry.csv")
montana_glycemic_registry$Date = mdy(montana_glycemic_registry$Date)
# looks identical
###################################

###### IMPORT UW DATA #############
uw_registry <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_registry.csv") #cffid and start
uw_bmi <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_bmi.csv") # date, cffid, height, bmi (height bmi need rename)
uw_pft <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_pft.csv") # datee, cffid, fev, fvc need renaming
uw_cgm <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_cgm.csv") # same code as montana
uw_demo <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_demo.csv") 
uw_genotypes <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_registry.csv") # no genotype data here?
uw_glycemic <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_glycemic.csv") # looks good, same as chco
uw_gtubes <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_gtubes.csv") # rename date of encounter to date
uw_pancreatic <- read_csv("./Prepost Trikafta/Data_Cleaned/uw_pancreatic.csv") # same as montana, no pancreatic data though

@@

###### IMPORT UM DATA ########
#um_all <- read_csv("./Prepost Trikafta/Data_Cleaned/um_all.csv")
###################################

@@

# keep original data, add in a new one to match CHCO
montana_demo = montana_demo %>% rename(race_original = Race)
montana_demo = montana_demo %>% mutate(Race = case_when(race_original == "White" ~ "Caucasian",
                                                        race_original == "Black or African American" ~ "Black",
                                                        race_original == NA ~ "Unknown",
                                                        TRUE ~ "Other"))
# rename the bmi_pft to chco names
# montana_bmi_pft = montana_bmi_pft %>% rename(FVC = `Measure of FVC`,
#                                              FEV1 = `Measure of FEV1`,
#                                              Height = `Patients height`,
#                                              Weight = `Weight of Patient`)
montana_bmi_pft$Date = mdy(montana_bmi_pft$Date)

@@

# keep original data, add in a new one to match CHCO
uw_demo = uw_demo %>% rename(race_original = Race)
uw_demo = uw_demo %>% mutate(Race = case_when(race_original == "White" ~ "Caucasian",
                                                        race_original == "Black or African American" ~ "Black",
                                                        race_original == NA ~ "Unknown",
                                                        TRUE ~ "Other"))
# make bmi pft
uw_bmi_pft = full_join(uw_bmi, uw_pft)
# rename the bmi_pft to chco names
# uw_bmi_pft = uw_bmi_pft %>% rename(FVC = `Measure of FVC`,
#                                    FEV1 = `Measure of FEV1`,
#                                    Height = `Patients height`,
#                                    BMI = `BMI Value`)

@@

# export csv analysis_data
write.csv(analysis_data, file = "./Prepost Trikafta/Data_Cleaned/analysis_data.csv")
```
